### 2022-05-22 r-journal/manuscript.tex

% !TeX root = RJwrapper.tex
\title{Partitioned Local Depth (PaLD) Clustering Analyses in R}
\author{by Lucy D'Agostino McGowan, Katherine Moore, and Kenneth Berenhaut}

\maketitle

\abstract{%
An abstract of less than 150 words.
}

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Partitioned Local Depths (PaLD) is a framework for a holistic
consideration of the community structure of distance-based data.
Leveraging a socially inspired perspective, the method provides
network-based community information which is founded on a new measure of
local depth and pairwise cohesion (partitioned local depth). The method
does not require distributional assumptions, optimization criteria, nor
extraneous inputs. A complete description of the perspective, together
with a discussion of the underlying social motivation, theoretical
results, and applications to additional data sets is provided in
\citet{berenhaut2022social}.

Building on existing approaches to (global) depth, local depth expresses
features of centrality as an interpretable probability which is free of
parameters and robust to outliers. Then, partitioning the probability
which defines local depth, we obtain a measure of cohesion between pairs
of points. Both local depth and cohesion reflect aspects of relative
position (rather than absolute distance) and provide a straightforward
way to account for varying density across the space. Specifically, as
shown in \citet{berenhaut2022social}, provided that two sets are
separated (in the sense that the minimum between-set distance is greater
than the maximum within-set distance), cohesion is invariant under the
contraction and dilation of the distances within each set. This property
may be particularly valuable when one has reason to believe that the
magnitude of distances varies across the space.

As cohesion captures a sense of the relationship strength between
points, we can then visualize the resulting community structure with a
network whose edges are weighted by (mutual) cohesion. The underlying
social framework motivates an elegant threshold for distinguishing
between strongly and weakly cohesive pairs.

As seen throughout this paper, we display the network obtained from
cohesion using a force-directed graph drawing algorithm and emphasize
the strong ties (colored by connected component). We refer to the
connected components of the network of strong ties as community
``clusters.'' Note that to qualify as a cluster in this definition, one
may not have any strong ties with those outside the cluster, and thus
the existence of disjoint groups is a strong signal for separation.
Here, clusters are identified without additional user inputs nor
optimization criteria. If one wishes to further break the community
graph into groups, one may use community detection methods for networks
(such as spectral clustering or the Louvain algorithm). Though only
briefly considered here, one may also use the collection of strong ties
in place of (weighted) k-nearest neighbors in settings such as
classification and smoothing. Overall, the structural information
obtained from local depth, cohesion and community graphs can provide a
holistic perspective on the data which does not require the use of
distributional assumptions, optimization criteria nor additional user
inputs.

We present a new package, \CRANpkg{pald}, for calculating Partitioned
Local Depths (PaLD) probabilities, implementing clustering analyses, and
creating data visualizations to display community structure. This paper
will describe how to use the package as well as walk through two
examples.

\hypertarget{pald}{%
\subsection{pald}\label{pald}}

The main functions in \CRANpkg{pald} package can be split into 3
categories:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A function for computing the cohesion matrix
\item
  Functions for extracting useful information from the cohesion matrix,
  such as local depths, neighbors, clusters, and graph objects
\item
  Plotting functions for community graphs
\end{enumerate}

In addition, the package provides a number of pertinent example data
sets commonly used to demonstrate cluster algorithms, including a
synthetic data set of two-dimensional points created by
\citet{gionis1clustering} to demonstrate clustering aggregation,
clustering data generated from the scikit-learn Python package
\citep{pedregosa2011scikit}, data describing cognate relationships
between words across 87 Indo-European languages \citep{dyen92}, data
compiled by \cite{tissue} of tissue gene expressions, and three example
data sets generated for the \citet{berenhaut2022social} paper.

While it is not a necessity, the \CRANpkg{pald} package is designed to
function well with the pipe operator, \texttt{\textbar{}\textgreater{}}.
This functionality will be demonstrated below.

\hypertarget{creating-the-cohesion-matrix}{%
\subsubsection{Creating the cohesion
matrix}\label{creating-the-cohesion-matrix}}

The input for the Partitioned Local Depths (PaLD) is a distance matrix
or \texttt{dist} object. Note that the collection of input distances (or
dissimilarities) does not need to satisfy the triangle inequality.

For demonstration purposes, we will show how one can compute a distance
matrix from an input data frame with, say, two variables \texttt{x1} and
\texttt{x2}. The input data may be of any dimension. The methods put
forth here work on data frames with higher dimensions, as described in
the \textbf{Examples} section; we are simply choosing a small data frame
here for demonstration purposes.

\begin{Schunk}
\begin{Sinput}
library(pald)
df <- data.frame(
  x1 = c(6, 8, 8, 16, 4, 14),
  x2 = c(5, 4, 10, 8, 4, 10)
)
rownames(df) <- c("A", "B", "C", "D", "E", "F")
\end{Sinput}
\end{Schunk}

The \texttt{dist()} function converts an input data frame into a
distance matrix, as demonstrated below. If the data are already provided
as a distance matrix (or \texttt{dist} object), the user can skip to the
next step.

\begin{Schunk}
\begin{Sinput}
d <- dist(df)
\end{Sinput}
\end{Schunk}

The function above creates a \texttt{dist} object. If converted to a
matrix, this will be a \(n\times n\) distance matrix, where \(n\)
corresponds to the number of observations in the original data frame, in
this example \(n = 6\).

This \texttt{dist} object, or a distance matrix, can then be passed to
the \texttt{cohesion\_matrix()} function in order to calculate the
pairwise cohesion values.

Cohesion is an interpretable probability that reflects the strength of
alignment of two points within local regions. It captures aspects of the
relative positioning of points and accounts for varying density across
the space.

\begin{Schunk}
\begin{Sinput}
d <- dist(df)
cohesion_matrix(d)
\end{Sinput}
\begin{Soutput}
#>            A          B          C         D          E         F
#> A 0.25000000 0.18333333 0.06666667 0.0000000 0.18333333 0.0000000
#> B 0.14000000 0.24000000 0.05000000 0.0000000 0.10666667 0.0000000
#> C 0.07333333 0.07333333 0.20333333 0.0000000 0.03333333 0.0800000
#> D 0.00000000 0.00000000 0.00000000 0.2333333 0.00000000 0.1333333
#> E 0.14000000 0.10666667 0.03333333 0.0000000 0.24000000 0.0000000
#> F 0.00000000 0.00000000 0.05000000 0.1400000 0.00000000 0.2400000
#> attr(,"class")
#> [1] "cohesion_matrix" "matrix"          "array"
\end{Soutput}
\end{Schunk}

Equivalently, the user can use the native pipe
\texttt{\textbar{}\textgreater{}} as follows.

\begin{Schunk}
\begin{Sinput}
df |>
  dist() |>
  cohesion_matrix()
\end{Sinput}
\begin{Soutput}
#>            A          B          C         D          E         F
#> A 0.25000000 0.18333333 0.06666667 0.0000000 0.18333333 0.0000000
#> B 0.14000000 0.24000000 0.05000000 0.0000000 0.10666667 0.0000000
#> C 0.07333333 0.07333333 0.20333333 0.0000000 0.03333333 0.0800000
#> D 0.00000000 0.00000000 0.00000000 0.2333333 0.00000000 0.1333333
#> E 0.14000000 0.10666667 0.03333333 0.0000000 0.24000000 0.0000000
#> F 0.00000000 0.00000000 0.05000000 0.1400000 0.00000000 0.2400000
#> attr(,"class")
#> [1] "cohesion_matrix" "matrix"          "array"
\end{Soutput}
\end{Schunk}

The \emph{cohesion matrix} output by the \texttt{cohesion\_matrix()}
function is the main input for the majority of the remaining functions.

\hypertarget{functions-for-extracting-information-from-the-cohesion-matrix}{%
\subsubsection{Functions for extracting information from the cohesion
matrix}\label{functions-for-extracting-information-from-the-cohesion-matrix}}

From the \emph{cohesion matrix}, a variety of useful quantities can be
calculated. Below, we create a cohesion matrix using the functions
described in the previous section.

\begin{Schunk}
\begin{Sinput}
df |>
  dist() |>
  cohesion_matrix() -> cohesion
\end{Sinput}
\end{Schunk}

The \texttt{local\_depths()} function calculates the \emph{depths} of
each point, outputting a vector of local depths. Local depth is an
interpretable probability which reflects aspects of relative position
and centrality via distance comparisons (i.e., \(d(z, x) < d(z, y)\)).

\begin{Schunk}
\begin{Sinput}
local_depths(cohesion)
\end{Sinput}
\begin{Soutput}
#>         A         B         C         D         E         F 
#> 0.6833333 0.5366667 0.4633333 0.3666667 0.5200000 0.4300000
\end{Soutput}
\end{Schunk}

In this case, the deepest point is \texttt{A}.

The \texttt{strong\_threshold()} function will calculate the cohesion
threshold for strong ties. This is equal to half the average of the
diagonal cohesion matrix.\citep{berenhaut2022social} This is a threshold
that may be used to distinguish between strong and weak ties.

\begin{Schunk}
\begin{Sinput}
strong_threshold(cohesion)
\end{Sinput}
\begin{Soutput}
#> [1] 0.1172222
\end{Soutput}
\end{Schunk}

In this case, the threshold is \texttt{0.117}.

The function \texttt{cohesion\_strong()} will update the cohesion matrix
to set all weak ties to zero (via the \texttt{strong\_threshold()}
function). Optionally, the matrix will also be symmetrized, using the
entry-wise (parallel) minimum of the cohesion matrix and its transpose,
with the default parameter \texttt{symmetric\ =\ TRUE}.

\begin{Schunk}
\begin{Sinput}
cohesion_strong(cohesion)
\end{Sinput}
\begin{Soutput}
#>      A    B         C         D    E         F
#> A 0.25 0.14 0.0000000 0.0000000 0.14 0.0000000
#> B 0.14 0.24 0.0000000 0.0000000 0.00 0.0000000
#> C 0.00 0.00 0.2033333 0.0000000 0.00 0.0000000
#> D 0.00 0.00 0.0000000 0.2333333 0.00 0.1333333
#> E 0.14 0.00 0.0000000 0.0000000 0.24 0.0000000
#> F 0.00 0.00 0.0000000 0.1333333 0.00 0.2400000
#> attr(,"class")
#> [1] "cohesion_matrix" "matrix"          "array"
\end{Soutput}
\end{Schunk}

The \texttt{community\_graphs()} function takes the cohesion matrix and
creates \CRANpkg{igraph} objects, graphs that describe the relationship
between the points. This function will output a list of three objects:

\begin{itemize}
\tightlist
\item
  \texttt{G}: the weighted (community) graph whose edge weights are
  mutual cohesion
\item
  \texttt{G\_strong}: the weighted (community) graph consisting of edges
  for which mutual cohesion is greater than the threshold for strong
  ties
\item
  \texttt{layout}: the graph layout, using the Fruchterman Reingold (FR)
  force-directed graph drawing for the graph \texttt{G}
\end{itemize}

\begin{Schunk}
\begin{Sinput}
graphs <- community_graphs(cohesion)
graphs[["G_strong"]]
\end{Sinput}
\begin{Soutput}
#> IGRAPH 8eef810 UNW- 6 3 -- 
#> + attr: name (v/c), weight (e/n)
#> + edges from 8eef810 (vertex names):
#> [1] A--B A--E D--F
\end{Soutput}
\end{Schunk}

Here we see that there are three connected components, ties \texttt{A-B}
and and \texttt{A-E} which form the first cluster, and the tie
\texttt{D-F} which forms another.

The \texttt{any\_isolated()} function will check whether there are any
isolated points that will inadvertently be dropped by a graph.

\begin{Schunk}
\begin{Sinput}
any_isolated(cohesion)
\end{Sinput}
\end{Schunk}

Here, there are no isolated points.

The ``clusters'' identified by PaLD are the connected components of the
graph of strong ties, \texttt{G\_strong}. To directly calculate them, we
can use the \texttt{community\_clusters()} function. This will output a
data frame with two columns, the first will correspond to the
\texttt{point}, as identified by the row name of the original input data
frame, \texttt{df}, the second will identify the \texttt{cluster} that
each point belongs to.

\begin{Schunk}
\begin{Sinput}
community_clusters(cohesion)
\end{Sinput}
\begin{Soutput}
#>   point cluster
#> A     A       1
#> B     B       1
#> C     C       2
#> D     D       3
#> E     E       1
#> F     F       3
\end{Soutput}
\end{Schunk}

In this example, three clusters are identified with these six points.
Points \texttt{A}, \texttt{B}, and \texttt{E} fall into cluster 1. Point
\texttt{C} is in cluster 2 and points \texttt{D} and \texttt{F} fall
into cluster 3.

\hypertarget{plotting-functions}{%
\subsection{Plotting functions}\label{plotting-functions}}

The final category of function is functions for data visualization. We
can begin by visualizing the points in data frame \texttt{df} (Figure
\ref{fig:fig1}). When visualizing these points, it is important to have
the aspect ratio of the x and y axes equal to 1 so as to not distort the
distances. When using the \CRANpkg{ggplot2} package for this
visualization, you can use the \texttt{coord\_fixed(ratio\ =\ 1)}
function. If using the \texttt{plot()} function included in the base
library, you can use the \texttt{asp\ =\ 1} argument.

\begin{Schunk}
\begin{Sinput}
library(ggplot2)
ggplot(df, aes(x1, x2)) +
  geom_text(label = rownames(df)) + 
  coord_fixed(ratio = 1) + 
  xlim(c(4, 16)) + 
  ylim(c(4, 16))
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig1-1} \caption[Visualize the points from data frame `df`]{Visualize the points from data frame `df`}\label{fig:fig1}
\end{figure}
\end{Schunk}

We can pass the cohesion matrix to the
\texttt{plot\_community\_graphs()} function to view the relationship
between points (Figure \ref{fig:fig2}).

\begin{Schunk}
\begin{Sinput}
plot_community_graphs(cohesion)
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
\end{figure}
\end{Schunk}

The \texttt{layout} argument allows the user to pass a matrix to dictate
the 2-dimensional layout of the graph. For example, if we wanted the
graph to match the visualization displayed in Figure \ref{fig:fig1}, we
can pass \texttt{as.matrix(df)}, or a matrix of the data frame
\texttt{df} to the \texttt{layout} argument (Figure \ref{fig:fig3}).
Additionally, this \texttt{plot\_community\_graphs()} function will also
permit parameters that can be passed to the \texttt{plot.igraph()}
function. For example, we can pass arguments to the \texttt{plot.igraph}
function via \texttt{...}, for example to increase the vertex size and
change the vertex label color, we can specify
\texttt{vertex.size\ =\ 100} and
\texttt{vertex.label.color\ =\ "white"}. Additionally, to allow axes, we
use \texttt{axes\ =\ TRUE} and to put them back on the original scale we
set \texttt{rescale\ =\ FALSE}, resetting the axis limits using
\texttt{xlim} and \texttt{ylim}. The \texttt{par(pty\ =\ "s")} function
forces the subsequent plot to be square.

\begin{Schunk}
\begin{Sinput}
par(pty = "s")

plot_community_graphs(cohesion, 
                      layout = as.matrix(df),
                      vertex.size = 100,
                      vertex.label.color = "white",
                      axes = TRUE,
                      rescale = FALSE,
                      asp = 1,
                      xlim = c(4, 16),
                      ylim = c(4, 16))
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
\end{figure}
\end{Schunk}

\hypertarget{examples}{%
\subsection{Examples}\label{examples}}

We will demonstrate the utility of the \CRANpkg{pald} package in three
clustering examples.

\hypertarget{clustering-tissue-gene-expression-data}{%
\subsubsection{Clustering tissue gene expression
data}\label{clustering-tissue-gene-expression-data}}

The first example is from a subset of tissue gene expression data from
\citet{zilliox2007gene}, \citet{mccall2011gene}, and
\citet{mccall2014gene}, obtained from the \textbf{tissuesGeneExpression}
bioconductor package \citep{tissue}. A \texttt{dist} object was created
using this data set and is included the \CRANpkg{pald} package in an
object called \texttt{tissue\_dist}.

The \texttt{tissue\_dist} object is a \texttt{dist} object resulting in
a distance matrix with 189 rows and 189 columns.

We can create the cohesion matrix using the \texttt{cohesion\_matrix}
function.

\begin{Schunk}
\begin{Sinput}
tissue_cohesion <- cohesion_matrix(tissue_dist)
\end{Sinput}
\end{Schunk}

The \texttt{community\_clusters()} function can be used to identify the
clusters of each tissue sample. Since the output is a data frame, we can
summarize the clusters using commonly used data analysis techniques. For
demonstration purposes, we will use the \CRANpkg{dplyr} package to
summarize the contribution of clusters.

\begin{Schunk}
\begin{Sinput}
community_clusters(tissue_cohesion) |>
  dplyr::count(cluster, point)
\end{Sinput}
\begin{Soutput}
#>    cluster       point  n
#> 1        1 endometrium 15
#> 2        1      kidney 39
#> 3        2 hippocampus 31
#> 4        3  cerebellum 26
#> 5        4  cerebellum  1
#> 6        5       colon 33
#> 7        6       colon  1
#> 8        7       liver  7
#> 9        8  cerebellum  1
#> 10       9       liver 17
#> 11      10  cerebellum  2
#> 12      11       liver  2
#> 13      12  cerebellum  1
#> 14      13  cerebellum  4
#> 15      14  cerebellum  2
#> 16      15  cerebellum  1
#> 17      16    placenta  2
#> 18      17    placenta  1
#> 19      18    placenta  3
\end{Soutput}
\end{Schunk}

From this, we can glean that cluster one consists of two types of
tissue, the kidney and endometrium. Cluster two is comprised of only the
hippocampus.

We can also display the relationships between tissue samples using the
\texttt{plot\_community\_graphs()} function (Figure \ref{fig:fig4}). For
clarity of the display, we show how to remove the labels using
\texttt{show\_labels\ =\ FALSE}. We will instead color by the labels by
passing these to the \texttt{vertex.color} parameter through the
\texttt{...} to the \texttt{plot.igraph} function. Similarly, we can add
a legend using the \texttt{legend()} function, as you would for an
\CRANpkg{igraph} visualization. Additionally, we use the
\texttt{edge\_width\_factor} and \texttt{emph\_strong} arguments to
adjust the width of the lines between and within PaLD clusters.

\begin{Schunk}
\begin{Sinput}
labels <- rownames(tissue_cohesion)
plot_community_graphs(tissue_cohesion,
                      show_labels = FALSE,
                      vertex.size = 4,
                      vertex.color = as.factor(labels),
                      edge_width_factor = 35,
                      emph_strong = 5) 
legend("topleft", 
       legend = unique(as.factor(labels)), 
       pt.bg = unique(as.factor(labels)),
       col = "black",
       pch = 21)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{figure}
\includegraphics[width=1\linewidth]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data. The line colors indicate the PaLD clusters, the point colors indicate the tissue classification.}\label{fig:fig4}
\end{figure}
\end{Schunk}

\hypertarget{cognate-based-language-families}{%
\subsection{Cognate-based Language
Families}\label{cognate-based-language-families}}

This example performs a PaLD analysis on a data set from \citet{dyen92}
that examines the relationship between 87 Indo-European languages from
the perspective of cognates. A \texttt{dist} object was created from
this data set and included in the \CRANpkg{pald} package in an object
called \texttt{cognate\_dist}.

This example will demonstrate how you can apply functions in the
\CRANpkg{igraph} package to objects output from the \CRANpkg{pald}
package. We can first use the \texttt{cohesion\_matrix()} function to
calculate the cohesion matrix and the \texttt{community\_graphs()}
function to create a list with the weighted community graph, the
weighted community graph with only strong ties included (and all others
set to 0), and the layout. From this, we can extract the graph with only
the strong ties, here called \texttt{cognate\_graph\_strong}.

\begin{Schunk}
\begin{Sinput}
cognate_cohesion <- cohesion_matrix(cognate_dist)
cognate_graphs <- community_graphs(cognate_cohesion)

cognate_graph_strong <- cognate_graphs[["G_strong"]]
\end{Sinput}
\end{Schunk}

We can then use the \texttt{neighbors()} function from the
\CRANpkg{igraph} package to extract the strong neighbors in this graph.
For example, if we wanted to extract all neighbors where the language is
``French'', we would run the following.

\begin{Schunk}
\begin{Sinput}
french_neighbors <- igraph::neighbors(cognate_graph_strong, "French")
french_neighbors
\end{Sinput}
\begin{Soutput}
#> + 8/87 vertices, named, from 9c17296:
#> [1] Italian         Ladin           Provencal       Walloon        
#> [5] French_Creole_C French_Creole_D Spanish         Catalan
\end{Soutput}
\end{Schunk}

Similarly, we can print the associated neighborhood weights by
subsetting the cohesion matrix.

\begin{Schunk}
\begin{Sinput}
cognate_cohesion["French", french_neighbors]
\end{Sinput}
\begin{Soutput}
#>         Italian           Ladin       Provencal         Walloon French_Creole_C 
#>      0.01997696      0.02094596      0.02871174      0.03258771      0.02406057 
#> French_Creole_D         Spanish         Catalan 
#>      0.02406057      0.01679733      0.01859688
\end{Soutput}
\end{Schunk}

\hypertarget{clustering-generated-data}{%
\subsubsection{Clustering generated
data}\label{clustering-generated-data}}

The \CRANpkg{pald} package includes three randomly generated data frames
corresponding to plots from \citet{berenhaut2022social}:

\begin{itemize}
\tightlist
\item
  \texttt{exdata1} is a data set consisting of 8 points to recreate
  Figure 1 in \citet{berenhaut2022social}
\item
  \texttt{exdata2} is a data set consisting of 16 points to recreate
  Figure 2 in \citet{berenhaut2022social}
\item
  \texttt{exdata3} is a data set consisting of 240 points to recreate
  Figure 4D in \citet{berenhaut2022social}
\end{itemize}

Here, we will demonstrate how to use \texttt{exdata3}. These points were
generated from bivariate normal distributions with varying means and
variances. There are eight ``true'' clusters.

We will demonstrate how we can compare PaLD to two clustering methods:
\emph{k}-means and hierarchical clustering. The code below calculates
the cohesion matrix (\texttt{exdata\_cohesion}) as well as the clusters
via PaLD (\texttt{exdata\_pald}), \emph{k}-means
(\texttt{exdata\_kmeans}) and hierarchical clustering using complete
linkage (\texttt{exdata\_hclust}).

\begin{Schunk}
\begin{Sinput}
exdata_cohesion <- exdata3 |>
  dist() |>
  cohesion_matrix()

exdata_pald <- community_clusters(exdata_cohesion)$cluster

exdata_kmeans <- kmeans(exdata3, 8)$cluster

exdata_hclust <- exdata3 |>
  dist() |>
  hclust() |>
  cutree(k = 8) 
\end{Sinput}
\end{Schunk}

We can compare this to the clustering generated by \emph{k}-means and
hierarchical clustering (Figure \ref{fig:fig5}).

\begin{Schunk}
\begin{Sinput}
par(mfrow = c(1, 3), pty = "s")
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_pald],
  xlab = "",
  ylab = "",
  main = "PaLD Clusters",
  asp = 1
)
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_kmeans],
  xlab = "",
  ylab = "",
  main = "K-Means Clusters (k = 8)",
  asp = 1
)
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_hclust],
  xlab = "",
  ylab = "",
  main = "Hiearchical Clusters (k = 8)",
  asp = 1
)
\end{Sinput}
\begin{figure}
\includegraphics{manuscript_files/figure-latex/fig5-1} \caption[PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al]{PaLD clustering of randomly generated example data (from Figure 4D from Berenhaut et al. (2022)) compared to k-means and hierarchical clustering with k = 8.}\label{fig:fig5}
\end{figure}
\end{Schunk}

Cohesion is particularly useful when considering data with varying local
density, see discussion in \citep{berenhaut2022social}. Note that the
PaLD algorithm is able to detect the eight natural groups within the
data without the use of any additional inputs (e.g., number of clusters)
nor optimization criteria. Despite providing the ``correct'' number of
clusters (i.e., \(k = 8\)) both \emph{k}-means and hierarchical
clustering did not give the desired result.

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

This paper introduces the \CRANpkg{pald} package, demonstrating its
utility for providing parameter-free clustering which can easily be
implemented for a variety of data sets.

\bibliography{RJreferences}

\address{%
Lucy D'Agostino McGowan\\
Wake Forest University\\%
Winston-Salem, NC\\ 27106\\
%
%
%
\href{mailto:mcgowald@wfu.edu}{\nolinkurl{mcgowald@wfu.edu}}%
}

\address{%
Katherine Moore\\
Wake Forest Unversity\\%
Winston-Salem, NC\\ 27106\\
%
%
%
\href{mailto:mooreke@wfu.edu}{\nolinkurl{mooreke@wfu.edu}}%
}

\address{%
Kenneth Berenhaut\\
Wake Forest University\\%
Winston-Salem, NC\\ 27106\\
%
%
%
\href{mailto:berenhks@wfu.edu}{\nolinkurl{berenhks@wfu.edu}}%
}